Standardized crypto-loans on\\ the Cardano blockchain

### 2021-02-05 papers/actus2021/article.tex
Completed introduction. Some tidying up of text and bibliography.

\bigskip\noindent
List of contributions:
\begin{itemize}
\item \textquotedblleft Marlowe Labs\textquotedblright{} editor on Playground
(Dmytro) 
\item ACTUS Utility functions (Dmytro, Quanterall) 
\item executable ACTUS spec for PAM (Dmytro, Alex) 
\item executable ACTUS spec for LAM (Wyo Hackathon) 
\item Library for generating ACTUS contracts (Dmytro) 
\item ACTUS-like examples on Playground (Dmytro, Alex, Pablo) 
\item Marlowe DSL enhancements for Actus: Cond, Mul (Dmytro)
\item ACTUS spec generator for Agda (Dmytro) 
\item Auto-refund static analysis (Pablo)
\item Verification with QC + SMT (Dmytro) 
\end{itemize}

@@

In this paper we explore ways in which contracts described in ACTUS (Algorithmic Contract Types Unified Standards) can be defined in Marlowe, Plutus and Haskell. Of course, Plutus or Haskell are able to express these contracts, but rendering them in Marlowe brings extra advantages. Marlowe is defined to provide a range of guarantees by design: a Marlowe contract will only make a finite number of interactions with its environment, and its lifetime can be read off from the code for a contract; moreover, when the contract terminates, any assets held by the contract will automatically be returned to the participants. None of these guarantees can be provided by a general purpose language.

@@

In implementing ACTUS we are able to provide further assurance in three other ways. First, we are able to use the declarative nature of Haskell to transliterate ACTUS formulae term-by-term into an executable form. Secondly, we are able to use random, property-based testing to validate the Haskell implementation against a Java one. Finally, we are able to automatically \emph{generate} ACTUS contracts in Marlowe from the terms -- i.e. parameters -- of the contracts; for a simple loan these would include the start and end dates of the loan and the amount loaned (the `principal'). The generated contracts use the executable spec in calculating values of cash flows in the Marlowe contracts.

Implementing ACTUS provides a suitable benchmark against which to assess Marlowe, and we illustrate how the implementation has led to the addition of a conditional expression construct to the language.

A note on notation: \texttt{typewriter} font will be used for Marlowe constructs
% SJT and fragments?
while \emph{math} font will be used for mathematical formulas and pseudo-code.
% Have I got this right?
% dk14 looks right to me - I didn't know how to highlight Marlowe constructs properly

% SJT: added this overview of the paper
% it is traditional to end the introduction like this,
% 'signposting' what is in the rest of the paper.
Section \ref{background} of the paper covers the relevant financial background together with the ACTUS financial standard. Section \ref{executable} builds an executable specification of ACTUS in Haskell, and this is used in Section \ref{generation} to generate  the Marlowe code for an ACTUS contract from the terms of the contract. Section \ref{tokenization} explains how tokens are used to represent ownership of roles in a running contract, and Section \ref{assurance} describes how we provide assurance that  contracts behave as they should. Section \ref{related} examines related work and Section \ref{conclusion} concludes.

@@

In this section we give a brief introduction to financial contracts, and loans in particular, and then describe the ACTUS financial standard.

@@

The Algorithmic Contract Types Unified Standards (ACTUS)~\cite{actus} define the logic embedded in legal agreements that eventually
turn the contract terms into actual cash flows, or more generally
business events. Most of its basic contract types represent
different variations of lending contracts.

@@

We provide assurance to users of our Marlowe ACTUS contracts in a number of ways, as described here.

@@

Unlike current mainstream DeFi lending approaches - Marlowe ACTUS-based
implementation tends to rely on trade-matching instead of pooling.
While asset pooling is proven to be superior to order-book based approaches
when it comes to automated market makers, using it for lending has
been shown to be more susceptible to attacks\cite{flash-loan}.

@@

Actus provides additional benefit of being regulatory friendly, ACTUS
foundation provides a set of tools allowing Monte-Carlo simulations
of ACTUS contracts.

@@

We are grateful to Finley and Kegan McIlwaine of the University of Wyoming for their implementation of the LAM contract in ActusLabs, and to colleagues at Quanterall\footnote{\url{https://quanterall.com/}} for their contribution of utility functions to our ACTUS codebase.

>>>>

@@

In this paper we explore ways in which contracts described in ACTUS (Algorithmic Contract Types Unified Standards) can be defined in the contract langauges Marlowe, Plutus and Haskell. Of course, Plutus or Haskell are able to express these contracts, but rendering them in Marlowe brings extra advantages. Marlowe is defined to provide a range of guarantees by design: a Marlowe contract will only make a finite number of interactions with its environment, and its lifetime can be read off from the code for a contract; moreover, when the contract terminates, any assets held by the contract will automatically be returned to the participants. None of these guarantees can be provided by a general purpose language.

@@

In implementing ACTUS on Cardano we are able to provide further assurance in three other ways. First, we are able to use the declarative nature of Haskell to transliterate ACTUS formulas term-by-term into an executable form in Haskell. Secondly, we are able to use random, property-based testing to validate the Haskell implementation against another written in Java. Finally, we are able to automatically \emph{generate} ACTUS contracts in Marlowe from the terms -- i.e. parameters -- of the contracts; for a simple loan these would include the start and end dates of the loan and the amount loaned (the `principal'). The generated contracts use the executable spec in calculating values of cash flows in the Marlowe contracts.

The contribution of our work is to show that that it is possible to implement financial contracts on blockchain in a way that \emph{multiple forms of assurance} are provided: from the language itself, from the static analysis, and from custom property verification. The development environment for Marlowe, the Marlowe Playground~\footnote{\url{https://alpha.marlowe.iohkdev.io/\#/}}, also provides a simulation environment for contracts for stepping forwards (and backwards) through contract execution, and thus allowing users to validate that contracts perform as they should. In addition, implementing ACTUS provides a suitable benchmark against which to assess the design of Marlowe; we illustrate how the implementation has led to the addition of a conditional expression construct to the language.

In the remainder of the paper, Section \ref{background} covers the relevant financial background, including the ACTUS financial standard. Section \ref{executable} builds an executable specification of ACTUS in Haskell, and this is used in Section \ref{generation} to generate  the Marlowe code for an ACTUS contract from the contract terms. Section \ref{tokenization} explains how tokens are used to represent ownership of roles in a running contract, and Section \ref{assurance} describes how we provide assurance that  contracts behave as they should. Section \ref{related} examines related work and Section \ref{conclusion} concludes.

A note on notation: \texttt{typewriter} font will be used for Marlowe constructs
while \emph{math} font will be used for mathematical formulas and pseudo-code.

@@

In this section we give a brief introduction to financial contracts, and to loans in particular, and then describe the ACTUS financial standard.

@@

The Algorithmic Contract Types Unified Standards (ACTUS)~\cite{actus} define the logic embedded in legal agreements that eventually
turn the contract terms into actual cash flows, or more generally
business events. Most of its basic contract types represent
different variations of lending contracts. 
% SJT moved from related work section.
ACTUS provides additional benefit of being regulatory friendly, and the ACTUS
foundation provides a set of tools allowing Monte-Carlo simulations
of ACTUS contracts.

@@

This section explains how we provide assurance to users of our Marlowe ACTUS contracts by means of property-based testing and SMT-based static analysis.

@@

Unlike current mainstream DeFi lending approaches - Marlowe ACTUS-based
implementation tends to rely on trade-matching instead of pooling.
While asset pooling is proven to be superior to order-book based approaches
when it comes to automated market makers, using it for lending has
been shown to be more susceptible to attacks~\cite{flash-loan}.

@@

@@

We are very grateful to Finley and Kegan McIlwaine of the University of Wyoming for their implementation of the LAM contract in ActusLabs, and to colleagues at Quanterall%\footnote{\url{https://quanterall.com/}} 
for their contributions to the Marlowe ACTUS codebase.
